Cell Decompositions for Rank Two Quiver Grassmannians

### 2018-02-21 Grassmannian_cells.tex
Intro

>>>>

A quiver Grassmannian is a projective variety attached to a fixed quiver representation which parametrize subrepresentations of a fixed dimension. Throughout the last years, quiver Grassmannians became of growing interest. On the one hand, this is due to the fact that the generating functions of the Euler characteristics of quiver Grassmannians of exceptional representations can be found as cluster variables. On the other hand, they are clearly interesting on their own as they reveal many properties of the representation and its geometry. Even if it turned out that every projective variety can be realized as a quiver Grassmannian - which is maybe not surprising as the classification problem of indecomposable quiver representations is wild - it turns out that very interesting phenomena arise when restricting to certain quivers or representations with certain properties. For instance quiver Grassmannians attached to exceptional representations  are smooth. For tame quivers of types $\tilde A$ or $\tilde D$, it is known that every quiver Grassmannian attached to an indecomposable representation admits a cell decomposition into affine spaces. It is also conjectured that this is true for exceptional representations, i.e. particularly for preprojective and preinjective representations.

There are basically two possible ways to find cell decompositions of quiver Grassmannians if they exist. One is to find a non-trivial torus action on the quiver Grassmannian under consideration. If it is additionally smooth, one can apply a result of Bialynicki-Birula which shows that the quiver Grassmannian decomposes into affine bundles over the fixed point components. In particular, this shows that the quiver Grassmannian has a cell decomposition if the fixed point components do have it.
Another method is to consider short exact sequences of quiver representations which induce a map between the quiver Grassmannians of the middle term and products of quiver Grassmannians of the outer terms. This map is also known as Caldero-Chapoton map. 
If this sequence has certain properties - e.g. (almost) split and generalizations - cell decompositions of quiver Grassmannians attached to the outer terms transfer to a decomposition of quiver Grassmannians of the middle term.

In this paper, we combine these two methods in order to show that every quiver Grassmannian attached to an exceptional representation of a generalized Kronecker quiver admits a cell decomposition into affine spaces. The proof also shows that this is true for so-called truncated preprojective representations which appear as certain quotients of preprojective representations. It turns out that these are precisely those representations which can be obtained from indecomposable representations of dimension $(d,1)$ when applying reflection functors. Actually, we can prove that quiver Grassmannians of truncated preprojective representations only depend on the dimension vector of the representation itself and on the dimension vector of the subrepresentations.

As a first step, we show that a torus action with a proper fixed point set can be defined on every quiver Grassmannian attached to representations of any quiver with multiple arrows or non-oriented cycles which can be lifted to the universal (abelian) covering quiver. These are precisely those cases where the universal covering quiver differs from the original quiver. This lifting property holds particularly for so-called tree modules which include all exceptional representations. The fixed point set consists precisely of those subrepresentations which can also be lifted to the universal (abelian) covering quiver. Actually, it turns out that each fixed point component is itself a quiver Grassmannians attached to the lifted representation.

The next step is to investigate under which conditions the fixed point components admit a cell decomposition. Here the Caldero-Chapoton map comes into play. In the case of the generalized Kronecker quiver, it turns out that a natural filtration of a fixed preprojective representation by preprojectives of smaller dimension transfers to the universal covering quiver. These filtration can be successively described by short exact sequences. The main advantage when passing to the universal covering is that the preprojective subrepresentations of the same dimension become orthogonal which rigidifies the situation in a sense. In the end, this machinery can be used to recursively build cell decompositions of all quiver Grassmannians of lifted (truncated) preprojective representations. As all the quiver Grassmannians of the (non-lifted) representation are smooth this gives rise to a cell decomposition of these quiver Grassmannians.

As a benefit (?) of this construction, we obtain a graph theoretical description of the non-empty cells. More precisely, with every (truncated) preprojective representation we can associate a so-called $2$-quiver. Basically, such a quiver is obtained from a usual quiver when adding so-called $2$-arrows which connect fixed subquivers. Now with every subset of the vertices we can associate a dimension vector. If this subset is also strong successor closed, which is in fact a condition to be checked easily, it corresponds to a cell and vice versa. The upshot is that we get a one-to-one correspondence between these subsets and compatible pairs of the respective Dyck path which leads to a geometric explanation for the Dyck path combinatorics appearing in the theory of quiver Grassmannians.

\sayT{references need to be included}
